Active learning for efficient systematic reviews -

### 2020-05-11 manuscript/drafts/manuscript.Rmd
last fixes

---
title: Manuscript drafts
author: "Gerbrich Ferdinands"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  bookdown::pdf_document2:
    number_sections: no
    toc: no
bibliography: asreview.bib
link-citations: yes
csl: systematic-reviews.csl
---

<!-- --- -->
<!-- title: Manuscript drafts -->
<!-- author: "Gerbrich Ferdinands" -->
<!-- date: '`r format(Sys.time(), "%d %B, %Y")`' -->
<!-- output: -->

<!--   bookdown::pdf_document2: -->

<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->
<!-- --- -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)

@@

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on datasets from original systematic reviews.")
```

Keywords: Human-Machine interaction, active learning, systematic reviews, text classification

#### Introduction 
<!-- This study aims to contribute to this growing area of research by exploring  -->

@@

#### Results
#### Discussion

\newpage

# Introduction
Systematic reviews are top of the bill in building evidence in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. 
<!-- PRISMA definition:  -->
<!-- > "A systematic review attempts to collate all relevant evidences that fits pre-specified eligibility criteria to answer a specific research -->
<!-- question. 
add that this is a systematic/rigorous process?-->

However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question. An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. 
<!-- Besides time-consuming, the process of title and abstract screening is prone to human error [@Wang2020]. -->
<!-- (14 to 77,910, trimmed mean 1,286).  -->
<!-- Need to be screened by multiple researchers.  --> 
Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent - such as in the context of the current COVID-19 crisis - it is almost impossible to provide a review that is both timely and comprehensive. To ensure a timely review, reducing workload in systematic reviews is essential.

With advances in Artificial Intelligence (AI), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications, i.e. 95% recall [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model reorders publications to be screened by their likeliness to be relevant. The model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process. Reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
~~Additionally, although outside the scope of the current study, active learning models have the potential to reduce the number of publications needed to screen when combined with some sort of stopping criterion [@Yu2019].~~
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010]. Active learning is when the model can iteratively improve its predictions by allowing the model to choose the data from which it can learn [@Settles2012]. Active learning has proven to be an efficient strategy in large datasets where labels are scarce, which makes title and abstract screening an ideal candidate for such models. When applied in screening prioritization, the reviewer screens publications that are presented by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge in selecting the next publication to be screened by the reviewer.

@@

Although the application of learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. First, whilst previous studies have evaluated models in many forms and shapes [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010], all models adopt the same classification technique. Even though a plethora of classification techniques exists [@Aggarwal2012], no study in this area has made a comparison between models adopting different classification techniques. 
<!-- whilst there exists a  of classification techniques for an active learning model to adopt, research to date has confined to models adopting one classification strategy [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010]. Models adopting other classification techniques have not been investigated.  -->
 <!-- evaluated techniques other than Support Vector Machines (SVM) -->
Second, the lack of model replication on reviews from varying research contexts makes it impossible to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the nature of the different vocabularies used. Therefore, it is of interest to evaluate model performance across different research contexts.

<!-- Taken together, the question remains how active learning models for screening prioritization perform across different classification techniques (1) and review contexts (2). Hence, additional evaluations of active learning models are required. The current study aims to address these issues by answering the following research questions:  -->
Taken together, evaluations of active learning models are required across different classification techniques (1) and review contexts (2).  The current study aims to address these issues by answering the following research questions:

  __RQ1__ What is the performance of active learning models across different classification techniques?

  __RQ2__ Does performance of active learning models differ across systematic reviews from different research areas?

<!-- Although important issues in the application of active learning models on title and abstract screening have been addressed prior studies -->

@@

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. Combining latest insights from this area, we propose seven different active learning models for the purpose of identifying relevant publications in systematic review datasets. The models were chosen to maximize the number of identified relevant publications, while minimizing the number of publications needed to screen. 
Working towards a general consensus in this emerging field, model performance was assessed by conducting a retrospective simulation on six systematic review datasets. Datasets were collected from the fields of medicine, software engineering, psychology, behavioural public administration, and virology to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of AI-assisted title and abstract screening in the field of systematic review.

\textcolor{red}{The remaining part of this paper is organized as follows. The Methods section elaborates on the setup of the current study, starting with a task description, followed the Technical Details section which will cover the workings of active learning models for study selection in systematic reviews on a conceptual level.
The results section reports ... The discusssion section summarises the findings, comments on them and  summarises the main findings, discusses 
discuss limitations, draw conclusion and}

@@

# Methods 
## Task Description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening them at the title and abstract level. In active learning for screening prioritization, the screening process proceeds as follows:

- Start with the set of all unlabeled (titles and) abstracts, $\mathcal{U}$.
- The reviewer provides a label for a few records $x \in \mathcal{U}$, creating an set of labeled titles and abstracts $\mathcal{L}$. The label can be either relevant $x_R$ or irrelevant $x_I$.
- The active learning cycle starts:

  1. A classifier is trained on the labeled titles and abstracts, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts labels for all unlabelled titles and abstracts, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant title and abstract $x^* \in \mathcal{U}$

  4. The model asks the reviewer to screen this title and abstract, $x^*_?$

  5. The reviewer screens the title and abstract and provides a label, $x^*_R$ or $x^*_I$

  6. The labeled title and abstract is added to the training data, $x_{R \texttt{ xor } I} \in \mathcal{L}$

@@

- In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabeled title and abstracts. The relevant titles and abstracts are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all title and abstracts have been labelled.

@@

<!-- The reviewer starts labeling some instances in $\mathcal{U}$, creating labeled data set $\mathcal{L}$. -->
<!-- This is the starting point for the active learning model. -->
<!-- The model trains a classifier who predicts ...  -->
<!-- The reviewer and the are interacting in the active learning cycle. -->
<!-- Based on previous labelling decisions by the reviewer, the model constantly reorders the remaining publications in the dataset. -->

<!-- # Algorithm 1 -->
<!-- Just a sample algorithmn -->
<!-- \begin{algorithm}[H] -->
<!-- \DontPrintSemicolon -->
<!-- \SetAlgoLined -->
<!-- \KwResult{Write here the result} -->
<!-- \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} -->
<!-- \Input{Write here the input} -->
<!-- \Output{Write here the output} -->
<!-- \BlankLine -->
<!-- \While{While condition}{ -->
<!--     instructions\; -->
<!--     \eIf{condition}{ -->
<!--         instructions1\; -->
<!--         instructions2\; -->
<!--     }{ -->
<!--         instructions3\; -->
<!--     } -->
<!-- } -->
<!-- \caption{While loop with If/Else condition} -->
<!-- \end{algorithm}  -->

<!-- meaning the reviewer the reviewer screens as little as possible while inclusion rate = high (not true but is nice statistic?) -->
<!-- The model is balancing two goals: obtain all relevant publications with as little as reviewing possible. -->
<!-- Active learning for screening prioritization denotes the scenario in which the reviewer is labeling references that are presented by a machine learning model. The machine learning model learns from the reviewers' decision and uses this knowledge in selecting the next reference that will be presented to the reviewer. -->

<!-- #### Assumptions  -->
<!-- 1) decisions of the original SR are **ground truth** (benchmark) (oracle) -->
<!-- study assumptions: -->
<!-- - simulation mode  -->
<!-- - assumptions (prior knowledge etc, oracle = reviewer). -->

@@

<!-- Selecting papers is a two-step process: abstract & fulltext screening -->

<!-- x -->

<!-- Now there are a few technical details. Many different versions of such algorithms exist. Many of such algorithms have been described in the active learning literature and have been applied in the systematic reviewing process.  Not exhaustive, but the algorithm can apply many different strategies to arrive at its predictions, which can be divided in following parameters: classifier, feature extraction strategy, balancing, query strateg y.  -->

<!-- The current section describes how common issues in identifying relevant publication can be dealt with.  -->
<!-- key components  -->

## Technical details
A more detailed account of the active learning models is given in the following section.
The structure and functions of the key components of the models will be introduced to clarify the choices made in the design of the current study.

### Classification
To make predictions on the unlabeled publications, a classifier is trained on features from the set of previously labeled publications. 
<!-- The classifier predicts class of a given unlabeled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst the performance of several classification techniques has been investigated in the AI-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2018; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

@@

### Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low (<10%) as only a fraction of the publications belong to the relevant class. This poses a problem for training a classifier as there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers are well-suited to separate data into classes, but not to correctly identify one class [@Wallace2010]. Therefore, the class imbalance problem causes the classifier to miss relevant publications. This is evident in the case of a systematic review dataset where only one percent of publications are relevant. A model would achieve 99% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.

Previous studies have addressed the class imbalance problem by rebalancing the training data in different ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalance the training set by Dynamic Supersampling (DS). DS decreases the number of irrelevant publications in the training data, whereas the number of relevant publications are increased (by copy) such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the training data is not fixed, but dynamically updated and depends on the size of the training data, the total number of publications, and the ratio between the total number of labeled publications.

### Word embeddings 
To predict publication class, the classifier uses information from the publications in the dataset. 
Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is called 'word embeddings'.

A classical example of word embeddings is a 'bag of words' (bow) representation. For each text in the data set, the number of occurrences of each word is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. The bag-of-words method is simplistic and will highly value often occurring but otherwise meaningless words such as "and". A more sophisticated approach is Term-Frequency Inverse Document Frequency (TF-IDF). TF-IDF circumvents this problem by adjusting the term frequency in a text with the inverse document frequency, the frequency of a given word in the entire data set [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2Vec (D2V), capable of grasping the relations between words by learning to predict the words in the texts [@Le2014].

Miwa et al. found that active learning was more difficult on data from the social sciences compared to data from the medical sciences and were able to link this difficulty to a natural difference in text complexity between these research areas [@Miwa2014]. As the study by Miwa et al. adopted a bow approach [@Miwa2014], we hypothesize that a more sophisticated word embedding strategy has the potential to bridge the performance gap between these research areas.

### Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. 
A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that will be presented next will be those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

@@

## Models 
The seven models consist of the components described in the Technical Details section, adopting four different classification techniques and two different word embeddings approaches:

First, four models combining every classifier with TF-IDF word embeddings were investigated:

>>>>

---
title: Active learning for efficient systematic reviews - 
subtitle: Evaluating models across research areas 
author: "Gerbrich Ferdinands"
header-includes:
    - \usepackage{setspace}\doublespacing
    - \usepackage[left]{lineno}
date: '`r format(Sys.time(), "%d %B, %Y")`'
output:
  word_document: default
  bookdown::pdf_document2:
    extra_dependencies: subfig, algorithm2e, float
    number_sections: no
    toc: no
bibliography: asreview.bib
link-citations: yes
csl: elsevier-vancouver.csl
---

<!-- Active learning for detecting relevant publications - evaluating models across systematic reviews from different research areas.  -->

<!-- Evaluating active learning models across research contexts.  -->

<!--  generalizability  -->
<!-- --- -->
<!-- title: Manuscript drafts -->
<!-- author: "Gerbrich Ferdinands" -->
<!-- date: '`r format(Sys.time(), "%d %B, %Y")`' -->
<!-- output: -->

<!--   bookdown::pdf_document2: -->
<!--     extra_dependencies: subfig, algorithm2e -->
<!--     number_sections: no -->
<!--     toc: no -->
<!--   word_document: default -->
<!-- bibliography: asreview.bib -->
<!-- link-citations: yes -->
<!-- csl: systematic-reviews.csl -->
<!-- --- -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, fig.ncol = 2, out.width = "50%", fig.align = "center", fig.pos = "H") 
options(tinytex.verbose = TRUE)

@@

```{r all tables, results = 'hide'}
table_nums <- captioner(prefix = "Table")
# table  on datasets statistics
table_nums(name = "datasets", caption = "Statistics on the test datasets obtained from six original systematic reviews.")
table_nums(name = "atd", caption = "ATD values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "wss", caption = "WSS@95 values ($\\bar x (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")
table_nums(name = "rrf", caption = "RRF@10 values ($\\bar x, (\\hat s)$) for all model-dataset combinations, and median (MAD) for all datasets.")

```
\newpage

\linenumbers
# Abstract
#### Background 
<!-- This study aims to contribute to this growing area of research by exploring  -->

@@

#### Results
#### Conclusions
#### Keywords
Machine learning, Active learning, Systematic Review, Study Selection, Text classification, Text representation,

<!-- Besides time-consuming, the process of title and abstract screening is prone to human error [@Wang2020]. -->
<!-- (14 to 77,910, trimmed mean 1,286).  -->
<!-- Need to be screened by multiple researchers.  -->

# Background
Systematic reviews are top of the bill in research. A systematic review brings together all studies relevant to answer a specific research question [@PRISMA-PGroup2015]. Systematic reviews inform practice and policy [@Gough2002] and are key in developing clinical guidelines [@Chalmers2007]. However, systematic reviews are costly because they involve the manual screening of thousands of titles and abstracts to identify publications relevant to answering the research question.

<!-- An experienced reviewer takes on average 30 seconds to screen one title and abstract, whereas an inexperienced reviewer takes even longer [@Wallace2010]. A typical review of 5000 publications would require over 40 hours of uninterrupted screening, while typically only a fraction of the publications is relevant (mean = 2.94%, IQR=2.5 @Borah2017]). -->
Conducting a systematic review typically requires over a year of work by a team of researchers [@Borah2017]. Nevertheless, systematic reviewers are often bound to a limited budget and timeframe. Currently, the demand for systematic reviews exceeds the available time and resources by far [@Lau2019]. Especially when the need for guidelines is urgent - such as in the context of the current COVID-19 crisis - it is almost impossible to provide a review that is both timely and comprehensive.

To ensure a timely review, reducing workload in systematic reviews is essential. With advances in Machine Learning (ML), there has been wide interest in tools to reduce workload in systematic reviews [@Harrison2020]. Various learning models have been proposed, aiming to predict whether a given publication is relevant or irrelevant to the systematic review. Findings suggest that such models potentially reduce workload with 30-70% at the cost of losing 5% of relevant publications, i.e. 95% recall [@OMara-Eves2015].
<!-- To date, several studies have begun to examine the use of learning algorithms for screening prioritization. -->

A well-established approach in increasing efficiency in title and abstract screening is screening prioritization [@Cohen2009; @Shemilt2014]. In screening prioritization, the learning model reorders publications to be screened by their likeliness to be relevant. The model presents the reviewer with the publications which are most likely to be relevant first, thereby expediting the process of finding all of the relevant publications. Such an approach allows for substantial time-savings in the screening process as the reviewer can decide to stop screening after a sufficient number of relevant publications have been retrieved [@Yu2019]. Moreover, reviewing relevant publications early facilitates a faster transition of those publications to the next steps in the review process [@Cohen2009]. 
<!-- Additionally, several studies report increasing efficiency beyond saving time [@OMara-Eves2015].  -->

<!-- assist the reviewer in the screening process -->
<!-- Several studies have proposed active learning models for screening prioritization -->
Recent studies have demonstrated the effectiveness of screening prioritization by means of active learning models [@Yu2018; @Yu2019; @Miwa2014; @Cormack2014; @Cormack2015; @Wallace2010; @Gates2018a]. With active learning, the machine learning model can iteratively improve its predictions on unlabelled data by allowing the model to select the records from which it wants to learn [@Settles2012]. The model queries these records to a human annotator who provides them with a label, from which the model then updates its predictions. The general assumption is that by letting the model select which records are labelled, the model can achieve higher accuracy while requiring the human annotator to label as few records as possible [@Settles2009]. Active learning has proven to be an efficient strategy in large unlabelled datasets where labels are expensive to obtain [@Settles2009], which makes the screening phase in systematic reviewing an ideal candidate for such models because labelling the typical large number of publications is very costly. When active learning is applied in the screening phase, the reviewer screens publications that are selected by an active learning model. Subsequently, the active learning model learns from the reviewers' decision ('relevant', 'irrelevant') and uses this knowledge to update its predictions and to select the next publication to be screened by the reviewer.

@@

<!-- , the complex nature of the field is making it difficult to draw overarching conclusions about best practice [@OMara-Eves2015]. -->

The application of active learning models in reducing workload of systematic reviews has been extensively studied [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010; @Gates2018a]. Whilst previous studies have evaluated active learning models in many forms and shapes [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010], all studies used the same classification technique, namely Support Vector Machine. Findings from outside the field of active learning show that different classification techniques can serve different needs in the retrieval of relevant publications, for example recall versus precision [@Kilicoglu2009; @Aphinyanaphongs2004]. Therefore, it is essential to evaluate different classification techniques in the context of active learning models.
<!-- no study in this area has made a comparison between models adopting different classification techniques. -->
<!-- Since a plethora of techniques exist [@Aggarwal2012], it would be of interest to incorporate such models to this area as well.  -->
Moreover, another component known to influence performance of the models is the way how the textual content of titles and abstracts are represented in a model, called the feature extraction strategy [@Le2014; @Zhang2011]. Previous studies all adopt the widely used 'bag of words' strategy [@Yu2018; @Yu2019; @Miwa2014; @Wallace2010]. It is of interest to evaluate models using this effective but rather simplistic approach by comparing them to models adopting a more sophisticated strategy, called 'Doc2vec' [@Le2014]. Lastly, previous studies have mainly focussed on reviews from a single scientific field, like medicine [@Wallace2010; @Gates2018a] and software engineering [@Yu2018, @Yu2019]. Model replications on reviews from varying research contexts are essential to draw conclusions about the general effectiveness of active learning models [@OMara-Eves2015; @Marshall2020]. As far as known to the authors, Miwa et al [@Miwa2014] were the only researchers to make a direct comparison between two systematic reviews from different research areas, namely the social and the medical sciences. They found that active learning was more difficult on data from the social sciences due to the nature of the different vocabularies used. Therefore, it is of interest to evaluate model performance across different research contexts. Taken together, evaluations of active learning models in the context of systematic reviewing are required (1) across different classification techniques, (2) feature extraction strategies, and (3) review contexts.  The current study aims to address these issues by answering the following research questions:

|    __RQ1__ What is the performance of several active learning models across different classification techniques?
|    __RQ2__ What is the performance of several active learning models across different feature extraction strategies? 
|    __RQ3__ Does the performance of active learning models differ across systematic reviews from different research areas?
<!-- is effective but rather simplistic as it is solely based on the relative frequency of words in the texts, the ordering of words in the texts cannot be taken into account.  -->
<!-- . yet rather simplistic as they cannot take into account the structure of the texts -->
<!-- the effect of adopting a different, more sophisticated strategies has not been investigated t -->
<!-- previous another model component worth exploring is the feature extraction strategy,  T It is worth evaluating strategies  -->
<!-- It  rather simplistic. Paragraph vector. As various feature extraction strategies exist, it is worth evaluating the different strategies as well. as mentioned by [@Yu2018] -->
<!-- many facets of models  -->
<!-- Previous studies adopt ...  -->
<!-- little attention has been paid to [@Yu2018] mention paragraph vector -->
<!-- How to represent  -->
<!-- the representation of the title an abstracts   -->
<!-- whilst there exists a  of classification techniques for an active learning model to adopt, research to date has confined to models adopting one classification strategy [@Yu2018; @Yu2018; @Miwa2014; @Wallace2010]. Models adopting other classification techniques have not been investigated.  -->
 <!-- evaluated techniques other than Support Vector Machines (SVM) -->
<!-- Although important issues in the application of active learning models on title and abstract screening have been addressed prior studies -->

@@

The purpose of the current paper is to increase the evidence base on active learning models for reducing workload in title and abstract screening in systematic reviews. We adopt four different classification techniques (Naive Bayes, Linear Regression, Support Vector Machine, and Random Forest) and two different feature extraction strategies (TF-IDF and Doc2vec) for the purpose of maximizing the number of identified relevant publications, while minimizing the number of publications needed to screen. Model performance was assessed by conducting a retrospective simulation on six systematic review datasets. Datasets were collected from the fields of medicine [@Cohen2006; @Appenzeller-Herzog2019], software engineering [@Yu2018], psychology [@vandeSchoot2017], behavioural public administration [@Nagtegaal2019] and virology [@Kwok2020] to assess generalizability of the models across research contexts. The models, datasets and simulations are implemented in a pipeline of active learning for screening prioritization, called $\texttt{ASReview}$ [@ASReview2020]. $\texttt{ASReview}$ is an open source and generic tool such that users can adapt and add modules as they like, encouraging fellow researchers to replicate findings from previous studies. All scripts and data used are openly published to facilitate usability and acceptability of ML-assisted title and abstract screening in the field of systematic review.

<!-- . -->

The remaining part of this paper is organized as follows. The Technical details section elaborates on the characteristics of active learning models for identifying relevant publications in the context of systematic reviews. The Simulation study section describes the study that was designed to answer the research questions. The findings of the simulation study are reported in the Results section. The implications of the findings in context of previous research are discussed in the Discussion section, followed by this study's main conclusions in the Conclusion section.

@@

# Technical details 
What follows is a more detailed account of the active learning models. The structure and functions of the key components of the models are introduced to clarify the choices made in the design of the current study.

## Task description
The screening process of a systematic review starts with all publications obtained in the search. The task is to identify which of these publications are relevant, by screening them at the title and abstract level. In active learning for screening prioritization, the screening process proceeds as follows:

<!-- Just a sample algorithmn -->
<!-- \begin{algorithm}[H] -->
<!-- \DontPrintSemicolon -->
<!-- \SetAlgoLined -->
<!-- \KwResult{Write here the result} -->
<!-- \SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output} -->
<!-- \Input{Write here the input} -->
<!-- \Output{Write here the output} -->
<!-- \BlankLine -->
<!-- \While{While condition}{ -->
<!--     instructions\; -->
<!--     \eIf{condition}{ -->
<!--         instructions1\; -->
<!--         instructions2\; -->
<!--     }{ -->
<!--         instructions3\; -->
<!--     } -->
<!-- } -->
<!-- \caption{While loop with If/Else condition} -->
<!-- \end{algorithm} -->

- Start with the set of all unlabeled records (titles and abstracts), $\mathcal{U}$.
- The reviewer provides a label for a few, for example 5-10, records $x \in \mathcal{U}$, creating an set of labeled records $\mathcal{L}$. The label can be either Relevant $\langle x, \texttt{R}\rangle$ or Irrelevant $\langle x, \texttt{I}\rangle$.
- The active learning cycle starts:

  1. A classifier, $C$, is trained on the labeled records $\mathcal{L}$, $C = \texttt{train}(\mathcal{L})$

  2. The classifier predicts relevancy scores for all unlabelled records $\mathcal{U}$, $C(\mathcal{U})$

  3. Based on the predictions by $C$, the model selects the most relevant record $x^* \in \mathcal{U}$

  4. The model queries the reviewer to screen this record, $\langle x^*, \texttt{?}\rangle$

  5. The reviewer screens the record and provides a label, $\langle x^*, \texttt{R}\rangle$ or $\langle x^*, \texttt{I}\rangle$

  6. The newly labeled record is added to the training data, such that $x \in \mathcal{L}$ and $x \notin \mathcal{U}$

@@

In this active learning cycle, the model can incrementally improve its predictions on the remaining unlabeled title and abstracts. The relevant titles and abstracts are identified as early in the process as possible. The reviewer and the model keep interacting until the reviewer decides to stop or until all records been labelled.

@@

<!-- The reviewer starts labeling some instances in $\mathcal{U}$, creating labeled data set $\mathcal{L}$. -->
<!-- This is the starting point for the active learning model. -->
<!-- The model trains a classifier who predicts ...  -->
<!-- The reviewer and the are interacting in the active learning cycle. -->
<!-- Based on previous labelling decisions by the reviewer, the model constantly reorders the remaining publications in the dataset. -->
<!-- meaning the reviewer the reviewer screens as little as possible while inclusion rate = high (not true but is nice statistic?) -->
<!-- The model is balancing two goals: obtain all relevant publications with as little as reviewing possible. -->
<!-- Active learning for screening prioritization denotes the scenario in which the reviewer is labeling references that are presented by a machine learning model. The machine learning model learns from the reviewers' decision and uses this knowledge in selecting the next reference that will be presented to the reviewer. -->
<!-- #### Assumptions  -->
<!-- 1) decisions of the original SR are **ground truth** (benchmark) (oracle) -->
<!-- study assumptions: -->
<!-- - simulation mode  -->
<!-- - assumptions (prior knowledge etc, oracle = reviewer). -->

@@

<!-- Selecting papers is a two-step process: abstract & fulltext screening -->
## Class imbalance problem
There are two classes in the dataset: relevant and irrelevant publications. Typically, the inclusion rate is low as only a fraction of the publications belong to the relevant class (2.94% [@Borah2017]). The class imbalance causes the classifier to miss relevant publications, because there are far fewer examples of relevant than irrelevant publications to train on [@OMara-Eves2015]. Moreover, classifiers can achieve high accuracy but still fail to identify any of the relevant publications [@Wallace2010]. This is evident in the case of a systematic review dataset where only three percent of publications are relevant. A model would achieve 97% accuracy when classifying all publications as irrelevant, even though none of the relevant papers would have been correctly identified.

Previous studies have addressed the class imbalance problem by rebalancing the training data in different ways [@OMara-Eves2015]. To decrease the class imbalance in the training data, the models in the current study rebalances the training set by Dynamic Resampling (DR). DR undersamples the number of irrelevant publications in the training data, whereas the number of relevant publications are oversampled such that the size of the training data remains the same. The ratio between relevant and irrelevant publications in the rebalanced training data is not fixed, but dynamically updated depending on the number of publications in the available training data, the number of publications in the total data, and the ratio between relevant and irrelevant publications in the available training data.

## Classification
To make predictions on the unlabeled publications, a classifier is trained on features from the set of previously labeled publications. 
<!-- The classifier predicts class of a given unlabeled publication.  -->
A technique widely used in classification tasks is the Support Vector Machine (SVM). SVMs separate the data into classes by finding a multidimensional hyperplane [@Tong2001; @Kremer2014]. SVMs have been proven to be effective in active learning models for screening prioritization [@Yu2018; Miwa2014]. Moreover, SVMs are the currently the only classifier implemented in ready-to-use software tools implementing active learning for screening prioritization (Abstrackr [@Wallace2012b], Colandr [@Cheng2018], FASTREAD [@Yu2018], Rayyan [@Ouzzani2016], and RobotAnalyst [@Przybyla2018]).

Whilst the performance of several classification techniques has been investigated in the ML-aided title-and-abstract screening field in general [@Kilicoglu2009; @Aphinyanaphongs2004], the relatively new subfield of active learning for screening prioritization has not yet studied the performance of classifiers other than SVMs [@Yu2018; @Yu2018; @Cormack2014; @Cormack2015; @Miwa2014; @Wallace2010]. 
<!-- Whilst the SVM is a well-known technique, also classifiers have been employed in class prediction tasks in systematic reviews.  -->
<!-- In fact, research in this area to date has not yet determined the performance of classifiers other than SVMs. -->
The current study aims to address this gap by exploring performance of three classifiers besides SVM:

@@

## Feature extraction
To predict publication class, the classifier uses information from the publications in the dataset. Examples of such information are titles and abstracts. However, a model cannot make predictions from the titles and abstracts as they are; their textual content needs to be represented numerically. The textual information needs to be mapped to feature vectors. This process of numerically representing textual content is referred to as 'feature extraction'.

A classical example of feature extraction is a 'bag of words' (bow) representation. For each text in the data set, the term frequency - the number of occurrences of each word - is stored. This leads to $n$ features, where $n$ is the number of distinct words in the texts [@scikit-learn]. A serious weakness of this method is that it highly values often occurring but otherwise meaningless words such as "the". A more sophisticated bow approach is Term-Frequency Inverse Document Frequency (TF-IDF), which circumvents this problem by adjusting the term frequency in a text with the inverse document frequency, the frequency of a given word in the entire data set [@Ramos2003]. A downside of TF-IDF and other bow methods is that they do not take into account the ordering of words, thereby ignoring semantics. An example of an approach that aims to overcome this weakness is Doc2vec (D2V). Doc2vec represents texts by a neural network, capable of grasping semantics by learning to predict the words in the texts [@Le2014].

<!-- Miwa et al. found that active learning was more difficult on data from the social sciences compared to data from the medical sciences and were able to link this difficulty to a natural difference in text complexity between these research areas [@Miwa2014]. As the study by Miwa et al. adopted a bow approach [@Miwa2014], we hypothesize that a more sophisticated feature extraction strategy has the potential to bridge the performance gap between these research areas.  -->

## Query strategy
The active learning model can adopt different strategies in selecting the next publication to be screened by the reviewer. A strategy mentioned before is selecting the publication with the highest probability of being relevant. In the active learning literature this is referred to as certainty-based active learning [@Settles2012]. Another well-known strategy is uncertainty-based active learning, where the instances that are presented next are those instances on which the model's classifications are the least certain, i.e. close to 0.5 probability [@Settles2012]. Traditionally, this strategy trains the most accurate model because the model can learn the most from instances it is uncertain about. However, a study comparing performance of both strategies in detecting relevant publications found that the accuracy gain of uncertainty-based screening was not significant [@Miwa2014].

@@

# Simulation study
The section below describes the simulation study that was carried out to answer the research questions.

## Set-up
<!-- The seven models consist of the components described in the Technical Details section, adopting four different classification techniques and two different feature extraction strategies:  -->

To address __RQ1__, four models combining every classifier with TF-IDF feature extraction were investigated:
